\LaTeX\ Author Guidelines for ICCV Proceedings

### 2019-06-25 chapters/supplementary.tex
Delete comments

\begin{algorithm}
  \caption{Implementation of \algname{}, generalized algorithm for signed graph partitioning}
\hspace*{\algorithmicindent} \textbf{Input:} $\mathcal{G}(V,E,w^+,w^-)$ with $N$ nodes and $M$ edges; boolean \texttt{{\color{blue}addCannotLinkConstraints}} \\
\hspace*{\algorithmicindent} \textbf{Output:} Final clustering \\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
    % \Procedure{GraphEdgeContr}{{\color{blue}bool \emph{addConstraints}}}
      % \State $\mathcal{G}'\gets \mathcal{G}(V,E^+ \cup E^-)$ \Comment{Initialize the contracted graph}
      \State $\tilde{\mathcal{G}}(\tilde{V},\tilde{E}) \gets \mathcal{G}(V,E,w^+,w^-)$  \Comment{Init. contracted graph}
      \State \texttt{UF} $\gets$ initUnionFind($V$) \Comment{Init. data structure representing clustering}
    %   \State PQ $\gets$ Sort $e\in E$ in descending order of $|w_e|$
        \State PQ.push$(|w_e|, e) \quad \forall e \in E $  \Comment{Init. priority queue in desc. order of $|w_e|=|w_e^+ - w_e^-|$, $\mathcal{O}(|E|)$}
        \State \texttt{canBeMerged}$[e] \gets$ \texttt{True} $\,\,\, \forall e\in E$ \Comment{Init. cannot-link constraints}
      % \State $E_\dagger \gets \{\}$ \Comment{Set of must-not-link edges}
    %   \State PQ.push$(e,  ) \quad \forall e \in E $  
    \State
      \While{PQ is \textbf{not} empty}
        \State $\tilde{w}, e_{uv} \gets $ PQ.popHighest() \Comment{$\mathcal{O}(\log |E|)$}
        \State \textbf{assert} \texttt{UF}.find($u$) $\neq$ \texttt{UF}.find($v$) \Comment{Edges in PQ always link nodes in different clusters}
        % \If{ $e_{uv} \notin E' $} 
        %     \State \textbf{continue}
        % \EndIf
        \If{({\color{ForestGreen}\textbf{$\tilde{w} > 0$}}) \textbf{and} \texttt{canBeMerged}$[e_{uv}]$}
        %   \State $u,v \gets u,v \in V' : $
        %   \State $S_u \gets S \in \Pi$ : $ u \in S$
        %   \State $S_v \gets S \in \Pi$ : $ v \in S$
          \State PQ, \texttt{canBeMerged}, $\tilde{E}$ $\gets$ \textsc{UpdateNeighbors}($u,v$)
        %   \State mergeDoubleEdges($u,v$) \Comment{Update PQ, $E_\dagger, \mathcal{G}'$}

        %   \State Update costs of double edges;
        %   \State Propagate constrained flags of double edges;

@@

        %   \State $ S_u \gets S_u \cup S_v$

@@

          % \For{every new double edge}
          %   \State Delete double edges
          %   \State Insert new one with updated cost
          % \EndFor

@@

    %   \State

@@

      % \State
    % \EndProcedure

@@

      % \State $\mathcal{G}'\gets \mathcal{G}(V,E^+ \cup E^-)$ \Comment{Initialize the contracted graph}

@@

        % \EndIf

@@

    %   % \State

@@

  % \caption{where $\alpha$ is the slowly growing inverse Ackerman function}

@@

        % \includegraphics[width=\textwidth,trim=0.1in 0.4in 0.2in 0.2in,clip]{./figs/edge_contraction.png} % left bottom right top

@@

% \captionof{figure}{ 
% {\small Example of edge contraction. First row: original graph $\mathcal{G}$; clustering $\Pi$ (gray shaded areas) with dashed edges on cut; cannot-link constraints (violet bars). Second row: contracted graph $\tilde{\mathcal{G}}_\Pi$. In step ii), edge $e_{uv}$ is contracted and node $v$ deleted from $\tilde{\mathcal{G}}_\Pi$. In step iii), double edges $e_{tu}$ and $e_{tv}$ resulting from edge contraction are replaced by single edge with updated interaction $f(\tilde{\cost}(e_{tu}), \tilde{\cost}(e_{tv}))$, see definitions in Tab.~\ref{tab:linkage-criteria}. }
% \label{fig:edge_contraction_and_contr_graph}  }

@@

            % \multicolumn{2}{c||}{}  &   & \multicolumn{2}{c}{\textbf{Signed graphs}}  \\        
            % \cmidrule(l{.15em}){4-5}
            % \cline{4-5}
            % \multicolumn{2}{c||}{} & \textbf{Unsigned graphs} &  \multicolumn{2}{c}{\thead{Add Cannot-Link Constraints:}} \\        

@@

    % \captionof{table}{}
    % \label{tab:linkage_and_update_rules}

@@

% During the agglomerative process, the interaction between adjacent clusters has to be properly updated and recomputed.  % given by the linkage criterion defined in Sec. \ref{sec:notation} and
% Given a graph $\mathcal{G}(V,E,\cost)$ and a clustering $\Pi$, we define the \emph{contracted graph} $\tilde{\mathcal{G}}_\Pi(\tilde{V}, \tilde{E}, \tilde{\cost})$, with $\tilde{V} \subseteq V$ such that node $u\in \tilde{V}$ represents the associated cluster $S_u \in \Pi$. Edges in $\tilde{E}$ are defined by adjacency-relationships between clusters and edge weights $\tilde{\cost}_{uv}$ represent inter-cluster interactions $\interact(S_u,S_v)$. 
\paragraph*{Update rules} During the agglomerative process, the interaction between adjacent clusters has to be properly updated and recomputed, as shown in Algorithm \ref{main_alg}.  % given by the linkage criterion defined in Sec. \ref{sec:notation} and
An efficient way of implementing these updates can be achieved by representing the agglomeration as a sequence of \emph{edge contractions} in the graph. Given a graph $\mathcal{G}(V,E,\cost)$ and a clustering $\Pi$, we define the associated \emph{contracted graph} $\tilde{\mathcal{G}}_\Pi(\tilde{V}, \tilde{E}, \tilde{\cost})$, such that there exists exactly one representative $|\tilde{V} \cap S| = 1$ for every cluster $S \in \Pi$ . Edges in $\tilde{E}$ represent adjacency-relationships between clusters 
and the signed edge weights $\tilde{\cost}_e$ are given by inter-cluster interactions $\tilde{\cost}(e_{uv})=\interact_{S_u,S_v}$. 
For the linkage criteria tested in this work, when two clusters $S_u$ and $S_v$ are merged, the interactions between the new cluster $S_u \cup S_v$ and each of its neighbors depend only on the previous interactions involving $S_u$ and $S_v$. Thus, we can recompute these interactions by using an \emph{update rule} $f$ that does not involve any loop over the edges of the original graph $\mathcal{G}$:
\begin{equation}
  \interact(S_u \cup S_v,S_t) = f\Big[ \interact(S_u,S_t), \interact(S_v,S_t) \Big] = f(\tilde{\cost}(e_{ut}), \tilde{\cost}(e_{vt})) %= \max \{ \tilde{\cost}(e_{ut}), \tilde{\cost}(e_{vt}) \}
\end{equation}
In Fig. \ref{fig:edge_contraction_and_contr_graph} we show an example of edge contraction and we list the update rules associated to the linkage criteria we introduced in Table \ref{tab:linkage-criteria}.
  % As an example, given the single-linkage criterion defined in Table , the interaction between $S_u \cup S_v$ and one of its neighbors $S_t$ is simply given by:
% All the update rules tested in this article are listed in Table \ref{tab:linkage-criteria}.

% \algname{} was implemented by using a standard agglomerative clustering algorithm based on a priority queue and a union
\paragraph{Implementation} As we show in Algorithm \ref{detailed_alg}, our implementation of \algname{} is based on an union-find data structure and a heap allowing deletion of its elements. The algorithm starts with each node assigned to its own cluster and sorts all edges $e\in E$ in a heap/priority queue (PQ) by their absolute weight $|\cost_e|=|w_e^+ - w_e^-|$ in descending order, so that the most attractive and the most repulsive interactions are processed first. It then iteratively pops one edge $e_{uv}$ from PQ and, depending on the priority $\tilde{\cost}_{uv}$, does the following: in case of attractive interaction $\tilde{\cost}_{uv}>0$, provided that $e_{uv}$ was not flagged as a cannot-link constraint, then merge the connected clusters, perform an edge contraction of $e_{uv}$ in $\tilde{\mathcal{G}}_\Pi$ and update the priorities of new double edges as explained in Fig. \ref{fig:edge_contraction_and_contr_graph}. 
% For every new pair of double edges in $\tilde{\mathcal{G}}_\Pi$, update their priorities according to one of the update rules listed in Table \ref{tab:linkage-criteria} together with their cannot-link relationships. 
If, on the other hand, the interaction is repulsive ($\tilde{\cost}_{uv}\leq 0$) and the option \texttt{addCannotLinkContraints} of Alg. \ref{detailed_alg} is \texttt{True}, then the edge $e_{uv}$ is flagged as cannot-link constraint.

\paragraph*{Complexity} In the main loop, the algorithm iterates over all edges, but the only iterations presenting a complexity different from $\mathcal{O}(1)$ are the ones involving a merge of two clusters, which are at most $N-1$. By using a union-find data structure (with path compression and union by rank) the time complexity of \texttt{merge}$(u, v)$ and \texttt{find}($u$) operations is $\mathcal{O}(\alpha(N))$, where $\alpha$ is the slowly growing inverse Ackerman function. The algorithm then iterates over the neighbors of the merged cluster (at most $N$) and updates/deletes values in the priority queue ($\mathcal{O}(\log |E|)$). Therefore, similarly to a heap-based implementation of hierarchical agglomerative clustering, our implementation of \algname{} has a complexity of $\mathcal{O}(N^2 \log N)$. In the worst case, when the graph is dense and $|E|=N^2$, the algorithm requires $\mathcal{O}(N^2)$ memory. Nevertheless, in our practical applications the graph is much sparser, so $\mathcal{O}(|E|)=\mathcal{O}(N)$. 
% and \UPDATE{in Fig. \ref{fig:runtime_plot} we show how the empirical runtime of \algname{} is much closer to $\mathcal{O}(N \log N)$ than $\mathcal{O}(N^2 \log N)$.} 
With a single-linkage, corresponding to the choice of the \emph{Maximum} update rule in our framework, the algorithm can be clearly implemented by using the more efficient Kruskal's Minimum Spanning Tree algorithm with complexity $\mathcal{O}(N \log N)$. 
Moreover, in the next section, we present an efficient implementation of \algname{} with \emph{Absolute Maximum} linkage that has empirical $\mathcal{O}(N \log N)$ complexity. %\TODO{Empirical complexity plot?}

\begin{algorithm}
  \caption{Mutex Watershed Algorithm proposed by \cite{wolf2018mutex}}
\hspace*{\algorithmicindent} \textbf{Input:} $\mathcal{G}(V,E,w^+,w^-)$ with $N$ nodes and $M$ edges \\
\hspace*{\algorithmicindent} \textbf{Output:} Final clustering \\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
      \State \texttt{UF} $\gets$ initUnionFind($V$) 
      % \State
      \For{$(u,v)=e\in E$ in descending order of $|w_e|=|w_e^+ - w_e^-|$}
        \If{\texttt{UF}.find($u$) $\neq$ \texttt{UF}.find($v$)} \Comment{Check if $u,v$ are already in the same cluster}
          \If{({\color{ForestGreen}\textbf{$w_e > 0$}}) \textbf{and} \texttt{canBeMerged}($u,v$)}  \Comment{Check for cannot-link constraints}
            \State \texttt{UF}.merge($u,v$) and inherit constraints of parent clusters
          \ElsIf{({\color{red}\textbf{$w_e \leq 0$}})}
            \State Add cannot-link constraints between parent clusters of $u,v$
          \EndIf
        \EndIf
      \EndFor
      \State
    %   \State
      \Return Final clustering given by union-find data structure \texttt{UF}
      % \State
    % \EndProcedure
  \end{algorithmic}
  \label{alg:mutex_watershed}
  % \caption{where $\alpha$ is the slowly growing inverse Ackerman function}
\end{algorithm}

@@

\begin{prop} \label{prop:equiv_MWS}
The Mutex Watershed Algorithm \ref{alg:mutex_watershed} (MWS) with empirical $\mathcal{O}(N \log N)$ complexity introduced by \cite{wolf2018mutex} returns the same final clustering given by the \algname{} Algorithm \ref{detailed_alg} with the use of cannot-link constraints and an Absolute Maximum update rule:
\begin{equation}\label{eq:def_abs_max}
f_{\mathrm{Abs.Max.}}(\tilde{\cost}_1,\tilde{\cost}_2) = \begin{cases} 
            \tilde{\cost}_1 & \text{if}\,\, |\tilde{\cost}_1|>|\tilde{\cost}_2|\\
            \tilde{\cost}_2 & \text{otherwise}
             \end{cases} 
\end{equation}
\end{prop}
\begin{proof}
Both algorithms sort edges in descending order of the absolute interactions $|w_e|$ and then iterate over all of them. The only difference is that MWS, after merging two clusters, does not update the interactions between the new cluster and its neighbors. 
However, since with an Abs. Max. linkage the interaction between clusters is simply given by the edge with highest absolute weight $|w_e|$, the order by which edges are iterated over in \algname{} is never updated. Thus, both algorithms perform precisely the same steps and return the same clustering.
\end{proof}
\begin{prop}
The \algname{} Algorithm \ref{detailed_alg} with the Absolute Maximum linkage defined in Eq. \ref{eq:def_abs_max} returns the same final clustering whether or not cannot-link constraints are enforced. 
\end{prop}
\begin{proof}
% In Proposition \ref{prop:equiv_MWS} we observed that \algname{} with Abs. Max. linkage never updates the order by which edges are iterated over
In the \algname{} Algorithm \ref{detailed_alg}, the clustering is updated only when two clusters are merged and the condition at line 9 is satisfied. 
We also observe that, in the unconstrained version of \algname{}, the predicate \texttt{canBeMerged} at line 9 can never be false because cannot-link constraints are never introduced at line 14.
Let us now contradict the initial hypothesis and assume by absurd that the constrained version of \algname{} introduces a cannot-link constraints between two clusters sharing a positive interaction $\tilde{w}>0$ and outputs a different clustering as compared to the unconstrained version. 
This can happen only in the situation shown in Fig. \ref{fig:abs_max_proof_example}, when two clusters $u$ and $v$ are merged together and share a common neighboring node $t$ having the following two properties: a) $u$ and $t$ are already constrained and share a repulsive interaction $w_{ut}\leq0$, b) $v$ and $t$ share an attractive interaction $w_{vt}>0$ that is higher in absolute value $|w_{vt}|>|w_{ut}|$. 
Then, according to Eq. \ref{eq:def_abs_max}, the new merged cluster $uv$ and $t$ are constrained and share a positive interaction. 
But this case can never happen, since if $|w_{vt}|>|w_{ut}|$ then clusters $v$ and $t$ are merged before clusters $u$ and $t$ are constrained.  
\end{proof}

@@

% \subsection{Multicut and correlation clustering } 
% \TODO{Omit completely?} For any clustering $\Pi$ of $\mathcal{G}$, we define as $E^{\mathrm{internal}}_\Pi$ the set of edges linking nodes in the same cluster, and as $E_\Pi^{\mathrm{external}}$ the complementary set of edges whose incident nodes belong to distinct clusters:
% \begin{equation}
% E_\Pi^{\mathrm{internal}} \equiv \{ e_{uv} \in E \,|\, \exists S \in \Pi : u \in S \, \text{and} \, v \in S \}, \qquad E^{\mathrm{external}}_\Pi \equiv E \setminus E^{\mathrm{internal}}_\Pi.
% \end{equation}
% % \begin{align}
% % E_\Pi^0 &= \{ e_{uv} \in E \,|\, \exists S \in \Pi : u \in S \, \text{and} \, v \in S \}, \\
% % E^1_\Pi &= E \setminus E^0_\Pi.
% % \end{align}
% The set of edges $E_\Pi^{\mathrm{external}}$ is known as the \emph{multicut} of $\mathcal{G}$ w.r.t. clustering $\Pi$. The instance of the NP-hard \emph{weighted correlation clustering/multicut problem} 
% % w.r.t. $\mathcal{G}(V,E,\cost)$ 
% corresponds to finding the partitioning that optimally balance attraction and repulsion in the graph and is given by the following binary integer program \cite{kappes2011globally,chopra1991multiway,andres2015lifting}:
% \begin{equation}\label{eq:multicut_obj}
% % \min_\Pi \texttt{MC}(\Pi) \equiv
%  \min_\Pi \sum_{e\in E} \cost_e x_e^\Pi,  \qquad \text{where} \quad x^\Pi_e = 
%  \begin{cases} 
%  1 & \text{if } e\in E^{\mathrm{external}}_\Pi \\
%  0 & \text{otherwise}.
%  \end{cases}
% \end{equation}
% In the next chapters we will use this objective as measure of how balanced a clustering is.

@@

\paragraph{Multi-step pipelines} Given the 2D WSDT superpixels, we build a 3D region-adjacency graph such that each node represents a superpixel. The weights of the edges connecting neighboring superpixels are computed by taking an average over both short- and long-range affinities connecting the two regions. We then convert the edge probabilities to signed weights using the logarithmic mapping defined in Eq. \ref{eq:mappings} and solve the multicut problem on this graph. For our experiments, we use the approximate Kernighan-Lin solver \cite{keuper2015efficient,kernighan1970efficient} (WSDT+MC). In some cases, the long-range affinities predicted by the CNN can connect two superpixels that are not direct-neighbors. Thus, in these cases we introduce additional \emph{lifted} edges in the graph and an instance of the lifted multicut problem (WSDT+LMC). This time, similarly to the methods mentioned in \cite{beier2016efficient}, we used a combination of approximate solvers consisting in GAEC and Kernighan-Lin. 
%and      Other state-of-the-art multi-step-pipelines for neuron-segmentation first find 2D superpixels using WSDT and then apply a graph partitioning algorithm: in our comparison, we include one pipeline using hierarchical agglomerative clustering with average linkage (avgHC) and one using an approximation of the Lifted Multicut Problem (LMC) \cite{beier2016efficient}.

\subsection{\algname{} on the full CREMI dataset} \label{sec:appendix_exps_full_cremi}
 \paragraph{Pre-merge processing} For the predictions on the full dataset from the CREMI challenge, we used  the padded volumes provided by the challenge. The crops on which we performed a prediction have a size of $1500\times1500\times127=2.86\cdot 10^8$ voxels or larger. Building a graph with $10^8$ nodes can easily incur a large use of memory, so we decided to perform a preprocessing step by initially merging some nodes together. Simply down-sampling the predictions of the CNN would have led to a loss of resolution and performances in the most difficult parts of the dataset. Thus, we decided to pre-merge the most connected components of the graph that would be anyway clustered during the first iterations of \algname{}. To do this, we used a simple approach: we generated a boundary probability map by taking for each voxel an average over affinities in all directions (both short- and long-range ones) and we run THRESH to find the connected components. With this approach, pixels are pre-clustered only when they are far away (in all directions) from all predicted boundaries. 
 To make sure that in this preprocessing step different neurons are never merged together by mistake, we intersected these segments given by the conservative THRESH with the segments given by WSDT. %\TODO{show figure?}
 By using this method, we initialized \algname{} with a reduced graph that represents a true over-segmentation. In our experiments, the use of this preprocessing method did not impact the final scores achieved by \algname{} and significantly reduced its runtime. On the full datasets, we used only 10 \% of the long-range connections in the pixel-graph, since adding all of them did not improve the scores and only made the simulations much slower and memory inefficient.

@@

\begin{table}[t]
\centering
    \footnotesize
% \begin{minipage}[T]{0.85\textwidth}
%     \centering
        \begin{tabular}{l|c|c|c|c|c}
          \algname{} linkage & \makecell{CREMI-Score\\(higher better)}  & \makecell{Rand-Score\\(higher better)} & \makecell{VI-merge\\(lower better)} & \makecell{VI-split\\(lower better)} & \makecell{Runtime\\(lower better)} \\ \midrule
          % Pipeline & method & \textsc{No} & \textsc{Yes} \\

>>>>

\begin{algorithm}
  \caption{Implementation of \algname{}, generalized algorithm for signed graph partitioning}
\hspace*{\algorithmicindent} \textbf{Input:} $\mathcal{G}(V,E,w^+,w^-)$ with $N$ nodes and $M$ edges; boolean \texttt{{\color{blue}addCannotLinkConstraints}} \\
\hspace*{\algorithmicindent} \textbf{Output:} Final clustering \\
  \hspace*{\algorithmicindent} 
      \State $\tilde{\mathcal{G}}(\tilde{V},\tilde{E}) \gets \mathcal{G}(V,E,w^+,w^-)$  \Comment{Init. contracted graph}
      \State \texttt{UF} $\gets$ initUnionFind($V$) \Comment{Init. data structure representing clustering}
        \State PQ.push$(|w_e|, e) \quad \forall e \in E $  \Comment{Init. priority queue in desc. order of $|w_e|=|w_e^+ - w_e^-|$, $\mathcal{O}(|E|)$}
        \State \texttt{canBeMerged}$[e] \gets$ \texttt{True} $\,\,\, \forall e\in E$ \Comment{Init. cannot-link constraints}
      \While{PQ is \textbf{not} empty}
        \State $\tilde{w}, e_{uv} \gets $ PQ.popHighest() \Comment{$\mathcal{O}(\log |E|)$}
        \State \textbf{assert} \texttt{UF}.find($u$) $\neq$ \texttt{UF}.find($v$) \Comment{Edges in PQ always link nodes in different clusters}
        \If{({\color{ForestGreen}\textbf{$\tilde{w} > 0$}}) \textbf{and} \texttt{canBeMerged}$[e_{uv}]$}
          \State PQ, \texttt{canBeMerged}, $\tilde{E}$ $\gets$ \textsc{UpdateNeighbors}($u,v$)

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

\paragraph*{Update rules} During the agglomerative process, the interaction between adjacent clusters has to be properly updated and recomputed, as shown in Algorithm \ref{main_alg}.  
An efficient way of implementing these updates can be achieved by representing the agglomeration as a sequence of \emph{edge contractions} in the graph. Given a graph $\mathcal{G}(V,E,\cost)$ and a clustering $\Pi$, we define the associated \emph{contracted graph} $\tilde{\mathcal{G}}_\Pi(\tilde{V}, \tilde{E}, \tilde{\cost})$, such that there exists exactly one representative $|\tilde{V} \cap S| = 1$ for every cluster $S \in \Pi$ . Edges in $\tilde{E}$ represent adjacency-relationships between clusters 
and the signed edge weights $\tilde{\cost}_e$ are given by inter-cluster interactions $\tilde{\cost}(e_{uv})=\interact_{S_u,S_v}$. 
For the linkage criteria tested in this work, when two clusters $S_u$ and $S_v$ are merged, the interactions between the new cluster $S_u \cup S_v$ and each of its neighbors depend only on the previous interactions involving $S_u$ and $S_v$. Thus, we can recompute these interactions by using an \emph{update rule} $f$ that does not involve any loop over the edges of the original graph $\mathcal{G}$:
\begin{equation}
  \interact(S_u \cup S_v,S_t) = f\Big[ \interact(S_u,S_t), \interact(S_v,S_t) \Big] = f(\tilde{\cost}(e_{ut}), \tilde{\cost}(e_{vt})) 
\end{equation}
In Fig. \ref{fig:edge_contraction_and_contr_graph} we show an example of edge contraction and we list the update rules associated to the linkage criteria we introduced in Table \ref{tab:linkage-criteria}.

\paragraph{Implementation} As we show in Algorithm \ref{detailed_alg}, our implementation of \algname{} is based on an union-find data structure and a heap allowing deletion of its elements. The algorithm starts with each node assigned to its own cluster and sorts all edges $e\in E$ in a heap/priority queue (PQ) by their absolute weight $|\cost_e|=|w_e^+ - w_e^-|$ in descending order, so that the most attractive and the most repulsive interactions are processed first. It then iteratively pops one edge $e_{uv}$ from PQ and, depending on the priority $\tilde{\cost}_{uv}$, does the following: in case of attractive interaction $\tilde{\cost}_{uv}>0$, provided that $e_{uv}$ was not flagged as a cannot-link constraint, then merge the connected clusters, perform an edge contraction of $e_{uv}$ in $\tilde{\mathcal{G}}_\Pi$ and update the priorities of new double edges as explained in Fig. \ref{fig:edge_contraction_and_contr_graph}. 
If, on the other hand, the interaction is repulsive ($\tilde{\cost}_{uv}\leq 0$) and the option \texttt{addCannotLinkContraints} of Alg. \ref{detailed_alg} is \texttt{True}, then the edge $e_{uv}$ is flagged as cannot-link constraint.

\paragraph*{Complexity} In the main loop, the algorithm iterates over all edges, but the only iterations presenting a complexity different from $\mathcal{O}(1)$ are the ones involving a merge of two clusters, which are at most $N-1$. By using a union-find data structure (with path compression and union by rank) the time complexity of \texttt{merge}$(u, v)$ and \texttt{find}($u$) operations is $\mathcal{O}(\alpha(N))$, where $\alpha$ is the slowly growing inverse Ackerman function. The algorithm then iterates over the neighbors of the merged cluster (at most $N$) and updates/deletes values in the priority queue ($\mathcal{O}(\log |E|)$). Therefore, similarly to a heap-based implementation of hierarchical agglomerative clustering, our implementation of \algname{} has a complexity of $\mathcal{O}(N^2 \log N)$. In the worst case, when the graph is dense and $|E|=N^2$, the algorithm requires $\mathcal{O}(N^2)$ memory. Nevertheless, in our practical applications the graph is much sparser, so $\mathcal{O}(|E|)=\mathcal{O}(N)$. 
With a single-linkage, corresponding to the choice of the \emph{Maximum} update rule in our framework, the algorithm can be clearly implemented by using the more efficient Kruskal's Minimum Spanning Tree algorithm with complexity $\mathcal{O}(N \log N)$. 
Moreover, in the next section, we present an efficient implementation of \algname{} with \emph{Absolute Maximum} linkage that has empirical $\mathcal{O}(N \log N)$ complexity. %\TODO{Empirical complexity plot?}

\begin{algorithm}
  \caption{Mutex Watershed Algorithm proposed by \cite{wolf2018mutex}}
\hspace*{\algorithmicindent} \textbf{Input:} $\mathcal{G}(V,E,w^+,w^-)$ with $N$ nodes and $M$ edges \\
\hspace*{\algorithmicindent} \textbf{Output:} Final clustering \\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
      \State \texttt{UF} $\gets$ initUnionFind($V$) 
      \For{$(u,v)=e\in E$ in descending order of $|w_e|=|w_e^+ - w_e^-|$}
        \If{\texttt{UF}.find($u$) $\neq$ \texttt{UF}.find($v$)} \Comment{Check if $u,v$ are already in the same cluster}
          \If{({\color{ForestGreen}\textbf{$w_e > 0$}}) \textbf{and} \texttt{canBeMerged}($u,v$)}  \Comment{Check for cannot-link constraints}
            \State \texttt{UF}.merge($u,v$) and inherit constraints of parent clusters
          \ElsIf{({\color{red}\textbf{$w_e \leq 0$}})}
            \State Add cannot-link constraints between parent clusters of $u,v$
          \EndIf
        \EndIf
      \EndFor
      \State
      \Return Final clustering given by union-find data structure \texttt{UF}
  \end{algorithmic}
  \label{alg:mutex_watershed}
\end{algorithm}

@@

\begin{prop} \label{prop:equiv_MWS}
The Mutex Watershed Algorithm \ref{alg:mutex_watershed} (MWS) with empirical $\mathcal{O}(N \log N)$ complexity introduced by \cite{wolf2018mutex} returns the same final clustering given by the \algname{} Algorithm \ref{detailed_alg} with the use of cannot-link constraints and an Absolute Maximum update rule:
\begin{equation}\label{eq:def_abs_max}
f_{\mathrm{Abs.Max.}}(\tilde{\cost}_1,\tilde{\cost}_2) = \begin{cases} 
            \tilde{\cost}_1 & \text{if}\,\, |\tilde{\cost}_1|>|\tilde{\cost}_2|\\
            \tilde{\cost}_2 & \text{otherwise}
             \end{cases} 
\end{equation}
\end{prop}
\begin{proof}
Both algorithms sort edges in descending order of the absolute interactions $|w_e|$ and then iterate over all of them. The only difference is that MWS, after merging two clusters, does not update the interactions between the new cluster and its neighbors. 
However, since with an Abs. Max. linkage the interaction between clusters is simply given by the edge with highest absolute weight $|w_e|$, the order by which edges are iterated over in \algname{} is never updated. Thus, both algorithms perform precisely the same steps and return the same clustering.
\end{proof}
\begin{prop} \label{prop:abs_max_cannot_link_property}
The \algname{} Algorithm \ref{detailed_alg} with the Absolute Maximum linkage defined in Eq. \ref{eq:def_abs_max} returns the same final clustering whether or not cannot-link constraints are enforced. 
\end{prop}
\begin{proof}
In the \algname{} Algorithm \ref{detailed_alg}, the clustering is updated only when two clusters are merged and the condition at line 9 is satisfied. 
We also observe that, in the unconstrained version of \algname{}, the predicate \texttt{canBeMerged} at line 9 can never be false because cannot-link constraints are never introduced at line 14.
Let us now contradict the initial hypothesis and assume by absurd that the constrained version of \algname{} introduces a cannot-link constraints between two clusters sharing a positive interaction $\tilde{w}>0$ and outputs a different clustering as compared to the unconstrained version. 
This can happen only in the situation shown in Fig. \ref{fig:abs_max_proof_example}, when two clusters $u$ and $v$ are merged together and share a common neighboring node $t$ having the following two properties: a) $u$ and $t$ are already constrained and share a repulsive interaction $w_{ut}\leq0$, b) $v$ and $t$ share an attractive interaction $w_{vt}>0$ that is higher in absolute value $|w_{vt}|>|w_{ut}|$. 
Then, according to Eq. \ref{eq:def_abs_max}, the new merged cluster $uv$ and $t$ are constrained and share a positive interaction. 
But this case can never happen, since if $|w_{vt}|>|w_{ut}|$ then clusters $v$ and $t$ are merged before clusters $u$ and $t$ are constrained.  
\end{proof}

@@

@@

\paragraph{Multi-step pipelines} Given the 2D WSDT superpixels, we build a 3D region-adjacency graph such that each node represents a superpixel. The weights of the edges connecting neighboring superpixels are computed by taking an average over both short- and long-range affinities connecting the two regions. We then convert the edge probabilities to signed weights using the logarithmic mapping defined in Eq. \ref{eq:mappings} and solve the multicut problem on this graph. For our experiments, we use the approximate Kernighan-Lin solver \cite{keuper2015efficient,kernighan1970efficient} (WSDT+MC). In some cases, the long-range affinities predicted by the CNN can connect two superpixels that are not direct-neighbors. Thus, in these cases we introduce additional \emph{lifted} edges in the graph and an instance of the lifted multicut problem (WSDT+LMC). This time, similarly to the methods mentioned in \cite{beier2016efficient}, we used a combination of approximate solvers consisting in GAEC and Kernighan-Lin.

\subsection{\algname{} on the full CREMI dataset} \label{sec:appendix_exps_full_cremi}
 \paragraph{Pre-merge processing} For the predictions on the full dataset from the CREMI challenge, we used  the padded volumes provided by the challenge. The crops on which we performed a prediction have a size of $1500\times1500\times127=2.86\cdot 10^8$ voxels or larger. Building a graph with $10^8$ nodes can easily incur a large use of memory, so we decided to perform a preprocessing step by initially merging some nodes together. Simply down-sampling the predictions of the CNN would have led to a loss of resolution and performances in the most difficult parts of the dataset. Thus, we decided to pre-merge the most connected components of the graph that would be anyway clustered during the first iterations of \algname{}. To do this, we used a simple approach: we generated a boundary probability map by taking for each voxel an average over affinities in all directions (both short- and long-range ones) and we run THRESH with a conservative threshold parameter to find the connected components. With this approach, pixels are pre-clustered only when they are far away (in all directions) from predicted boundaries. 
 To make sure that in this preprocessing step different neurons are never merged together by mistake, we intersected the segments given by THRESH with the segments given by WSDT. 
 In our experiments, the use of this preprocessing method did not introduce any major changes in the final segmentations and scores achieved, but significantly reduced the runtime. Moreover, to reduce the runtime and memory requirements even further, we used only 10 \% of the long-range connections in the pixel-graph, since adding all of them did not improve the scores.

@@

\begin{table}[t]
\centering
    \footnotesize
        \begin{tabular}{l|c|c|c|c|c}
          \algname{} linkage & \makecell{CREMI-Score\\(higher better)}  & \makecell{Rand-Score\\(higher better)} & \makecell{VI-merge\\(lower better)} & \makecell{VI-split\\(lower better)} & \makecell{Runtime\\(lower better)} \\ \midrule
Average & \textbf{0.226} & \textbf{0.936} & 0.315 & 0.494 & 3.49 $\cdot$ 10$^4$ \\
Sum + CLC \cite{levinkov2017comparative} & 0.282 & 0.906 & 0.358 & 0.510 & 4.64 $\cdot$ 10$^4$ \\
Abs Max \cite{wolf2018mutex} & 0.322 & 0.897 & 0.286 & 0.735 & 1.24 $\cdot$ 10$^4$ \\
Max + CLC & 0.324 & 0.893 & 0.292 & 0.698 & 6.31 $\cdot$ 10$^4$ \\
Sum \cite{keuper2015efficient} & 0.334 & 0.872 & 0.461 & 0.444 & 4.74 $\cdot$ 10$^4$ \\
Average + CLC & 0.563 & 0.772 & 0.259  & 1.142 & 2.95 $\cdot$ 10$^4$ \\
Min & 2.522 & 0.030 & \textbf{0.197} & 6.365 & 2.97 $\cdot$ 10$^3$ \\
Min + CLC & 2.522 & 0.030 & \textbf{0.197}  & 6.365 & 4.77 $\cdot$ 10$^3$ \\
Max & 2.626 & 0.028 & 7.069 & \textbf{0.026} & \textbf{6.04 $\cdot$ 10$^\mathbf{2}$} \\
        \end{tabular}
        \vspace*{1.1em}
    \caption{Performances achieved by different versions of \algname{} on the CREMI 2016 training set. 
    CREMI-Score \cite{cremiChallenge}, is given by a combination of the Adapted Rand-Score (Rand-Score) and the Variation of Information Score for under-clustering (VI-merge) and over-clustering (VI-split) \cite{arganda2015crowdsourcing}.
    CLC stands for cannot-link constraints. For all algorithms, the chosen value of bias parameter was $\beta = 0$. We used a machine with CPU Intel(R) Xeon(R) X5650  @ 2.67GHz for our comparison experiments.}
    \label{tab:extended_results_cremi}
\end{table}

\subsection{\algname{} sensitivity to noise: adding artifacts to CNN predictions} \label{sec:appendix_noise_gen}
Additionally to the comparison on the full training dataset, we performed more experiments on a crop of the more challenging CREMI training sample B, where we perturbed the predictions of the CNN with noise and we introduced additional artifacts like missing or fictitious boundary evidences.
